---
title: "1_Getting_Started"
author: "Matt Steele"
date: "2022-07-08"
output: html_document
---

## Special Topic | Data Analysis and Visualization with R | Part I ----

  Matt Steele 
  Government Information Librarian
  West Virginia University 

=========================================================================

## Resources------------------------------------------------------------

    R download and documentation - https://cran.r-project.org/
    RStudio download and documentation - https://rstudio.com
    
    This workshop was developed using the O'Reilly Platform Lesson video
      R Programming for Statistics and Data Science
      https://libwvu.on.worldcat.org/oclc/1062397089

==========================================================================

## Setting Working Directory -------------------------------------------

The working directory is just a file path on your computer that sets the default location of any files you read into R, or save out of R. In other words, a working directory is like a little flag somewhere on your computer which is tied to a specific analysis project.


```{r warning= FALSE, message= FALSE}

# get working directory

  getwd()

# set working directory
  
  setwd()
  
# you can set wd by using menu > session > set working directory > choose directory

```

===============================================================================

## Importing Data ----------------------------------------------------------
    
    The read.csv function imports .csv files
    
```{r}

  help("read.csv")

```
    
    There several options possible. 
    If you have Tidyverse loaded you can also use the the read_csv.

There are additional built-in functions for reading data that is not in CSV format
      
```{r}

  help("read.table")

```


==================================================================
  ## Let's Try to load the dataset "coalpublic2020.csv" in our RScript
==================================================================

```{r}
  
  coalProd <-  read.csv("coalpublic2020.csv")
  coalProd

```


    In RStudio, you can use View(gss)
    
```{r}

    View(coalProd)

```
    
    
    
## Viewing Data and Variables ----------------------------------------------
    
```{r}

    help("names")
    names(coalProd)

```
    
    The names() function can display the variable names in the data.


```{r}

  help("head")
  head(coalProd)

```
    
    The head() function to see the first six rows of the dataset.
  

```{r}

  help("str")
  str(coalProd)

```
    
    The str() function allows you to see the structure of the dataset.
  

```{r}

  help("summary")
  summary(coalProd)

```
    
    The summary() function allows you to see descriptive statistics of the dataset.



==================================================================
  ## Let's try to us at the names(), head(), str(), and summary() functions in our RScript
==================================================================  
    
## Subsetting variables in data frames --------------------------------------
    
    You can use the $ sign to specify a variable within the data for analysis and you can run specific fun
    
```{r}

    head(coalProd$Mine.Name)

```
  
    
```{r}

    str(coalProd$Mine.Name)

```

## Saving and Exporting R ---------------------------------------------------
    
    
    Command history
    
```{r}

    history() #Last 25 commands
    history(max.show=Inf) #All commands during session
    savehistory(file="IntroR")
    getwd()

```
    
    
    You can use write.csv() to export a csv file of your data.
    
    
```{r}

    write.csv(coalProd, "coal_production_20202.csv")

```
    
