---
title: "Tidyverse_Cleaning_Manipulating"
author: "Matt Steele"
date: "2022-07-08"
output: html_document
---

## Special Topic | Data Analysis and Visualization with R | Part II ----

  Matt Steele 
  Government Information Librarian
  West Virginia University 
  
===========================================================================


## Packages ---------------------------------------------------------
    

```{r}

help("install.packages")

```

R is open source code after it's initial development people began adding to it with packages. An R Package is something that you can plug into RStudio to extend the basic functionality that is built in with R. One of the reasons that R has become so popular is because it has this rich ecosystem of packages that really make R a comprehensive platform for data science.
    
Packages have their own built-in functions that cannot be used unless the package has been loaded.
    
For this workshop we are going to use a package called Tidyverse that was created for people working with data. Let's load it up and install it now.

===========================================================================

## Let's try to install Tidyverse in our RScript

===========================================================================



## Loading Packages -----------------------------------------------------------

```{r}

help("library")
    
```

    You only have to install a package once on you computer
    
    However you do need to call a package, every session you plan on using it


===========================================================================

## Let's try to load  Tidyverse in our RScript

===========================================================================


## The Tidyverse Pacakge ----------------------------------------------------
  
    Collection of packages for data analysis and data visualizations
  
      ggplot2 - data visualization
        
```{r}

          browseURL("https://ggplot2.tidyverse.org/")

```
      
      tibble - lighter and more user-friendly version of data.frame()
     
             
```{r}

          browseURL("https://tibble.tidyverse.org/")

```

  
      tidyr - create tidy and meaningfully arranged data
      
              
```{r}

          browseURL("https://tidyr.tidyverse.org/")

```

  
      readr - better importation of data into R 

        
```{r}

          browseURL("https://readr.tidyverse.org/")

```

  
      purrr - better functional programming

        
```{r}

          browseURL("https://purrr.tidyverse.org/")

```

  
      dplyr - data manipulation tools
  
  
        
```{r}

          browseURL("https://dplyr.tidyverse.org/")

```
  
        
============================================================================

## Tibble Package ---------------------------------------------------------------

    Tibble allows you load a lighter and more user-friendly version of data.frame()
    
```{r}

  help("as_tibble")
  browseURL("https://tibble.tidyverse.org/")

  coalProd <- as_tibble(coalProd)

```

===========================================================================

## Let's try to use the as_tibble to make our data frame more user-friendly in our RScript

===========================================================================

## Readr Package ---------------------------------------------------------

    The readr package allows for a better importation of data and also allows you to import Excel and Google Sheets files
    
```{r}


  help("read_delim") #OR#
  help("read_csv")
  browseURL("https://readr.tidyverse.org/")
  
 
```


===========================================================================

## Let's try to use read_csv function to read a data frame about bigfoot sightings in our RScript
    https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv

===========================================================================



##### HELPFUL HINT ##########

    Because there can be crossover in the naming conventions of functions from packages, you can tell R specifically what package you want to call the function from using 
    
    package_name::function_name
    
```{r}

   bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')


```

==============================================================================

## Dplyr Package ---------------------------------------------------------------
  
    Dplyr allows you to use it's built-in functions to manipulate data
    
```{r}

    help("dplyr")
    browseURL("https://dplyr.tidyverse.org/")

```


============================================================================


## Manipulating Data --------------------------------------------------------


# Remove or re-code absent data
  
    We can remove absent data using the mutate function from Dplyr

    Remove NA values
  
```{r}

help("na.omit")

    view(bigfoot)

    bigfoot_clean <- na.omit(bigfoot)
    view(bigfoot_clean)

```
 
 
===========================================================================

## Let's remove all of the NAs from the bigfoot data frame in our RScript

===========================================================================


# Function: filter() - subsets data according to criteria
    
```{r}

?dplyr::filter

filter(coalProd, Mine.State == "West Virginia")


```


##### HELPFUL HINT ##########

    R accepts Boolean operators which will allow you to do things like filter more specifically
    
    |     is    OR
    &     is    AND
    ==    is    EQUAL
    >     is    LESS THAN
    <     is    GREATER THAN
    >=    is    LESS THAN OR EQUAL TO
    <=    is    GREATER THAN OR EQUAL TO
    !=    is    NOT EQUAL TO
    

===========================================================================

## Let's filter our Coal Production data frame to just show West Virginia Mines in Greenbrier county
and get the mean() of the Production in our RScript

===========================================================================    
    

```{r}

coalProd_wv_gb <- filter(coalProd, Mine.State == "West Virginia" & Mine.County == "Greenbrier")
  coalProd_wv_gb
  mean(coalProd_wv_gb$Production.Tons)

```

    # Function: select() - keeps ONLY selected variables
    
```{r}

?dplyr::select

coalProd_names <- select(coalProd, Mine.Name, Mine.State, Mine.Type)
coalProd_names

```

    # Function: mutate() - create a new variable and append it to the data frame
    
```{r}

?dplyr::mutate

bigfoot_clean <- mutate(bigfoot_clean, temperature_low_celcius = (temperature_low - 32) * 5/9)
view(bigfoot_clean)

coalProd <- mutate(coalProd, Prod.per.employee = Production.Tons/Average.Employees)
view(coalProd)

```

===========================================================================

## Let's create a new variable in our RScritpt that determines that amount of production 
per number of employees in our RScript

===========================================================================   

===================================================================================

##### HELPFUL HINT ##########

    If your created variable name is the same as an existing variable in the data frame, the created varaible's will override the observations of the existing variable
    
=====================================================================================


    # Function: transmute() - creates a new variable and discards the rest of the data frame
    
```{r}

?dplyr::transmute

```

```{r}

coalProd_per_employee <- transmute(coalProd, Prod.per.employee = Production.Tons/Average.Employees) 
  head(coalProd.per.employee)

```

    # Function: arrange() - sort data frame by variables
    
```{r}

?dplyr::arrange

```
    
```{r}

  arrange(coalProd, desc(Mine.State))
  arrange(coalProd, (Mine.State), desc(Mine.Status))

```

===================================================================================

##### HELPFUL HINT ##########

    THE PIPE OPERATOR ( %>% ) is an operator in dyplr that allows you to string together multiple functions to act on one specific object
    
=====================================================================================
   
    # Function: summarise() - get a summary of a group
    
```{r}

?dplyr::summarise

dplyr::summarise(coalProd, avgProd = mean(Production.Tons)) 

mean(coalProd$Production.Tons)

```
    
    # Function: group_by() - group a categorical variable and create a new table
 

```{r}

?dplyr::group_by


coalProd_byMineType <- coalProd %>% group_by(Mine.Type) %>% summarise(avgProd = mean(Production.Tons))
coalProd_byMineType


```
    
===========================================================================

## Let's create a data frame that calculates the average lwo temperature by state of bigfoot sightings

=========================================================================== 



    # Function: sample() - randomly sample from the data frame
    
```{r}

?dplyr::sample_n

?dplyr::sample_frac

coalProd_sample <- coalProd %>% sample_frac(.25, replace = T)

coalProd_sample <- coalProd %>% sample_frac(.25, replace = T) %>% group_by(Mine.Type) %>% summarise(avgProd = mean(Production.Tons))


```

===============================================================================

# Re-coding variables  --------------------------------------------------
  
    Adding a numeric value to an categorical variable
  
```{r}
  
    summary(coalProd$Mine.Type) # Use the summary or str function to see a variable's value label 
    coalProd$Mine.Type <- as.factor(coalProd$Mine.Type)
    
    coalProd <- mutate(coalProd, Mine.Type_num = recode_factor(Mine.Type, "Refuse" = 1, "Surface" = 2, "Underground" = 3, .ordered = F))
    
    view(coalProd) # see the result

```
 
    Re-coding a numerical variable into a categorical variable using mutate() and cut() functions
    
```{r}

help("cut")

```
    

```{r}

  summary(coalProd$Average.Employees)

  coalProd <- mutate(coalProd, Avg_Empl_bin = cut(Average.Employees, breaks=c(-Inf, 15, 500, Inf), labels=(c("less than 15", "15 to 500", "over 500"))))
  
    coalProd <- mutate(coalProd, Avg_Empl_bin_num = cut(Average.Employees, breaks=c(-Inf, 15, 500, Inf), labels=(c(1, 2, 3))))

  view(coalProd)

```


    Let's recode all NA values in a data frame as 999
    
```{r}

bigfoot <- bigfoot %>% mutate(title = ifelse(is.na(title), 999, title))

view(bigfoot)

bigfoot_filter <- filter(bigfoot, title != "999")
bigfoot_filter

```
    
    
    


# Descriptive variables ---------------------------------------------------------
  
    # Summary () will give you descriptive statistics for numeric and/ or frequency for factors and characters 

```{r}

summary(bigfoot_clean) # full data frame

summary(coalProd_wv_gb$Production.Tons) # subset variable

mean(coalProd_wv_gb$Production.Tons) # average of observed cases
median(coalProd_wv_gb$Production.Tons) # the middle number(s) of observed cases

min(coalProd_wv_gb$Production.Tons) # lowest number in observed cases
max(coalProd_wv_gb$Production.Tons)  # the middle number(s) of observed cases
range(coalProd_wv_gb$Production.Tons) # most commonly occurring number of observed cases

quantile(coalProd_wv_gb$Production.Tons, .25) #1st
quantile(coalProd_wv_gb$Production.Tons, .75) #2nd
fivenum(coalProd_wv_gb$Production.Tons) # vector of length 5 with the minimum, 25th percentile, median, 75th percentile, maximum values

sd(coalProd_wv_gb$Production.Tons) # standard deviation

 coEffVar_prod <-   sd(coalProd_wv_gb$Production.Tons) /  mean(coalProd_wv_gb$Production.Tons)
  coEffVar_prod # coefficient of variation

```


=============================================================================

## Exporting your data frame ---------------------------------------------


```{r}

help("write.csv")
help("write_csv")

coalProd_save <- write_csv(coalProd_wv_gb, "coal_production_wv_greenbrier_2020.csv")

```

=============================================================================

Let's try to export a data frame as a csv in our RScript file

=============================================================================



