---
title: "Visualize with GGplot2"
author: "Matt Steele"
date: "2022-09-17"
output: html_document
---



===========================================================================

## Geometrics of GGPlot2 

## How plot ----------------------------------------------------------------

    Basic steps:
      set the data (main function)
      choose a shape layer
      map variables to aesthetics using aes() -> start with x and y in main function
  
    Structure of Plots
      Data - data you plotting
      Aesthetics - what you map on x and y axis
      Geometries - how your data will be represented visually

# First three are mandatory to create a plot. The following four are for flourish:

    Facets - discrete subplots that you graph can be split into 
    Stats - statistical transformations you may choose to use
    Coordinates - where data is plotted
    Themes - non-data related information (fonts, colors)



## Visualize your data with GGplot2

=============================================================================

## Part I:  Bar Graph of Coal Production in WV

```{r}

##If you did not earlier, let's create a data frame that only selects WV mines in Greenbrier county


coalProd_wv_gb <- filter(coalProd, Mine.State == "West Virginia" & Mine.County == "Greenbrier")
  view(coalProd_wv_gb)
  

```

      # Next we have to create parameters using varaible. This first step will tell R that we are creating graph using the categorical variable Mine.
      
        aes - sets the X axis and y axis variables
      
      
```{r}

help("geom_col")


bar.coalProd <- ggplot(coalProd_wv_gb, aes(x = Company.Type, y = Production.Tons))
  bar.coalProd

```

      # Plot the data
      
```{r}

bar.coalProd + 
  geom_col() +
  theme_classic()
  

```

      ## Add some color and labels
      
```{r}

bar.coalProd + 
  geom_col(fill = "skyblue") +
  theme_classic() +
  labs(title = "Coal production by type of company in Greenbrier county",
      subtitle = "2020", 
      x = "Type of Mine", 
      y = "In tons")

```

============================================================================

Let's create a bar graph of the number of sighting using geom_bar() and export it using the bigfoot data frame

==========================================================================

## Part II: Line Graph

      # Let's create a line graph of cases over time and let's add some extra features to our plot
      
      # Begin by rading in our data fream
      
```{r}

# Create an object that contains elements that your would like to plot
  
  bidenApproval <- read.csv("approval_topline_biden_2022.csv", stringsAsFactors = T)
  bidenApproval <- as_tibble(bidenApproval)
  bidenApproval 

```

       # Clean the data frame
       
```{r}

    str(bidenApproval)
  
  #Fix Date
  
    bidenApproval$modeldate <- as.Date(bidenApproval$modeldate, tryFormats = c("%m/%d/%Y"))
    bidenApproval
  
  #Fix name of president variable 
  
    bidenApproval <- mutate(bidenApproval, president = recode_factor(president, "Joe Biden" = "Joe Biden", 'Joseph R. Biden Jr.' = "Joe Biden" ))
    str(bidenApproval)
  
```

        # Create a line graph using approval rates by date
      
```{r}

## Setting the Parameters with Aesthetics aes(x = ... , y = ...) --------------------------------------------------    

  bidenApproval.scatter <- ggplot(bidenApproval, aes(modeldate, approve_estimate))
  bidenApproval.scatter
  
  bidenApproval.scatter <- ggplot(bidenApproval, aes(x = modeldate, y = approve_estimate))
  bidenApproval.scatter
  
  
```

    ## Setting the type of plot with Geometries


```{r}


  bidenApproval.scatter + geom_point()

```

    ## Color by factor levels

```{r}
  
  bidenApproval.scatter + geom_point(aes(color = subgroup)) + 
    scale_color_manual(values=c("darkseagreen", "blueviolet", "coral")) +
    theme_minimal()
  
  
  #Save the factor color
  
  bidenApproval.scatter <- ggplot(bidenApproval, aes(x = modeldate, y = approve_estimate, color = subgroup, shape = subgroup))

```

      ## Add a reference line


```{r}

 #Add a Reference Line
  
  bidenApproval.scatter + geom_point() + 
    geom_hline(yintercept = 50, size = 3, color = "grey") +
    theme_minimal()
  
  # The Order of the grammar of plotting is important
  
  bidenApproval.scatter + 
    geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point() +
    theme_bw()
  
```

    ## Subsetting Factors with Facets

```{r}

  bidenApproval.scatter + 
    geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point() + facet_grid(subgroup ~.) + 
    scale_color_manual(values=c("darkseagreen", "blueviolet", "coral")) +
    theme_minimal()

```

    ## Add a Statistical Transformation


```{r}

  # Line to test for overplotting

  bidenApproval.scatter + geom_point() + facet_grid(subgroup ~.) + stat_smooth()
  
  # You can generally use geoms and stats interchangeably 
  
  bidenApproval.scatter + geom_point() + geom_smooth()
  
  # For instance you can use either to create a box plot stat_boxplot() == geom_boxplot()
  
```

    ## Theme() allows you to change the background of the plot

```{r}
  
  bidenApproval.scatter + 
    geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point() + 
    theme_update()
  
  
  bidenApproval.scatter + geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point() + 
    theme_classic()

```

    ## Color and Fill allows you to color variable

```{r}

  bidenApproval.scatter <- ggplot(bidenApproval, aes(x = modeldate))
  
  bidenApproval.scatter + 
    geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point (aes(y = approve_estimate), color = "darkseagreen") + 
    geom_point(aes(y = disapprove_estimate), color = "coral") + 
    theme_bw()
    
```

    ## Labs allow you to re-label x, y, title, and legend of the plot


```{r}

  bidenApproval.scatter + 
    geom_hline(yintercept = 50, size = 25, color = "grey", alpha = .25) + 
    geom_point (aes(y = approve_estimate), color = "darkseagreen") + 
    geom_point (aes(y = disapprove_estimate), color = "coral") +  
    labs(y="Percent Approval", 
         x = "Date", 
         title = "Joe Biden Presidential Approval", 
         color = "category") +
    theme_classic()

```

==========================================================================

## Part III: Histogram

      # Let's create a histogram of the bigfoot sighting during a waning moon in summer
      
```{r}

bigfoot_clean

# Let's limit the data frame to that mooon phases over .50 during the summer

bigfoot_waning <- filter(bigfoot_clean, moon_phase > .50 & season == "Summer")
bigfoot_waning

```

      # Set the parameters for looking at low temperatures for sighting
      
```{r}

bigfoot.hist <- ggplot(bigfoot_waning, aes(x = temperature_low))

```

      # Let's create a histogram looking at the distribution of the sightings during these specifications

```{r}

bigfoot.hist + 
  geom_histogram(color = "yellow", fill = "skyblue") + 
  theme_classic()

# What if we want to view a more categorized version using bins

bigfoot.hist + 
  geom_histogram(bins = 3, color = "yellow", fill = "skyblue") + 
  theme_classic()

# Let's change the parameter to use the variable numbers using bin width


bigfoot.hist + 
  geom_histogram(binwidth = 10 , color = "yellow", fill = "skyblue") + 
  theme_classic() +
  labs(title = "Distubution of Low Temperature", 
       caption = "low temperature of bigfoot sightings during a waning moon in summer",
       x = NULL,
       y = NULL)

```

===============================================================================

## Part IV: Exporting an image ----------------------------------------------

    # You can also export an image in RScript by using the Plots tab
  
```{r}

ggsave("temperature_low.png", height = 10, width = 10, dpi = 320) # saves the most recent plot

```


